---
title: "Data624 HW Week 1"
author: "Leo Yi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: no
    theme: paper
    highlight: tango
    font-family: Consolas
  pdf_document:
    toc: yes
---

```{=html}
<style type="text/css">

code {
  font-family: "Consolas";
  font-size: 11px;
}

pre {
  font-family: "Consolas";
  font-size: 11px;
}

mark {
  background-color: whitesmoke;
  color: black;
}

</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3.5)

library(fpp2)
library(seasonal)
```

<font size="5">Time Series and Decomposition</font>

### 2.1

Use the help function to explore what the series <mark> gold </mark>, <mark> woolyrnq </mark> and <mark> gas </mark> represent

```{r}
# ?gold
# ?woolyrnq
# ?gas
```

#### a - autoplot

Use <mark> autoplot() </mark> to plot each of these in separate plots.

```{r}
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

#### b - frequency

What is the frequency of each series? Hint: apply the <mark> frequency() </mark> function.

```{r}
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

#### c - outlier

Use <mark> which.max() </mark> to spot the outlier in the <mark> gold </mark> series. Which observation was it?

```{r}
max_val_time <- which.max(gold)
max_val <- gold[max_val_time]

cat('At time', max_val_time, 'the max value was', max_val)
```

---

### 2.3

Download some monthly Australian retail data from [the book website](https://otexts.com/fpp2/extrafiles/retail.xlsx). These represent retail sales in various categories for different Australian state, and are stored in a MS-Excel File.

#### a - Import

You can read the data into R with the following script:
  
    retaildata <- readxl::read_excel("retail.xlsx", skip = 1) 
    
The second argument ( **skip=1** ) is required becasue the Excel sheet has two header rows.
    
```{r}    
path <- 'C:/Downloads/docs/data624/week1/'
filename <- 'retail.xlsx'
filepath <- paste0(path, filename)

retaildata <- readxl::read_excel(filepath, skip = 1)
```
  
#### b - Select Time Series

Select one of the time series as follows (but replace the column name with your own chosen column):
  
    myts <- ts(retaildata[,"A3349873A"],
      frequency=12, start=c(1982,4))
      
```{r}
myts <- ts(retaildata[,101], frequency=12, start=c(1982,4))
```

#### c - Explore Time Series

Explore your chosen retail time series using athe following functions:
  
    autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()
    
Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

```{r}
autoplot(myts)
```

There seems to be clear seasonality and an upward trend, as well as a possible cycle that starts around 2010.

```{r}
ggseasonplot(myts)
```

Here, the seasonality is very apparent, with retail sales peaking in December, likely due to holiday sales.

```{r}
ggsubseriesplot(myts)
```

Again, December is clearly a leader in sales compared to other months. Each month over the years follows a trending pattern with growth in the beginning, leveling off, and then another period of growth.

```{r, fig.height = 8}
gglagplot(myts)
```

Here, we see almost all months showing a positive correlations, with lag 12 having the strongest association. 

```{r}
ggAcf(myts)
```

The autocorrelation chart here shows peaks every 12 months pointing at the seasonality of December sales, as well as significant associate of sales for neighboring periods.

---

### 6.2

The <mark>plastics</mark> data set consists of monthly sales (in thousands) of product A for a plastics manufacturer for five years.

#### a - plot time series

Plot the time series of sales of Product A. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r}
autoplot(plastics)
```

There seems to be an upward trend over the 5 year period, as well as a seasonal high just after halfway through each year.

#### b - classical multiplicative decomposition

Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r, fig.height = 7}
plastic_m_decomposed <- decompose(plastics, type = 'multiplicative')

autoplot(plastic_m_decomposed)

# plastics %>%
#   decompose(type = 'multiplicative') %>%
#   autoplot()
```

#### c - plot evaluation

Do the results support the graphical interpretation from part a?

  - It looks like we can see the upward trend mentioned earlier, but it also looks like the trend stops at the end of the 5 year period
  - The seasonality is clearer in the decomposition, appearing to be waves that dip around the start of January and tops out around July/August.

#### d - seasonally adjusted data

Compute and plot the seasonally adjusted data.

```{r}
plastics_s_adj <- plastics / plastic_m_decomposed$seasonal

autoplot(plastics_s_adj) +
  ggtitle('Plastic Sales, Seasonally Adjusted') +
  ylab('Sales')
```

#### e - introduction of outlier

Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
plastics[14]

(plastics[14] <- plastics[14] + 500)

(plastics / decompose(plastics, type = 'multiplicative')$seasonal) %>%
  autoplot

```

Inflating record 14, which is two months after the first year, and recalculating the seasonally adjusted data shows a clear outlier when visualized in a plot.

#### f - timing consideration

Does it make any difference if the outlier is near the end rather than in the middle of the time series?

  - There would be a difference at the ends since we have less data to calculate trends and seasonality.


